My Title

### 2025-12-09 misc/collapse cheat sheet/collapse_cheat_sheet.Rnw
Replace sebkrantz.github.io/collapse -> fastverse.github.io/collapse.

Calling \code{help("collapse-documentation")} brings up a detailed documentation, which is also available \href{https://sebkrantz.github.io/collapse/reference/index.html}{online}. See also the \href{https://fastverse.github.io/fastverse/}{\emph{fastverse}} package/project for a recommended set of complimentary packages and easy package management.
%}

@@

\textbf{Examples}
<<>>=
mtcars |> fsubset(mpg > fnth(mpg, 0.95), disp:wt, cylinders = cyl)
mtcars |> colorder(cyl, vs, am, pos = 'after') |> head(2)
i <- base::invisible # These are equivalent, the second option is faster:
mtcars |> fgroup_by(cyl, vs, am) |> fmutate(sum_mpg = fsum(mpg)) |> i()
mtcars |> fmutate(sum_mpg = fsum(mpg, list(cyl, vs, am), TRA = 1)) |> i()
# These are also equivalent (weighted means), again the second is faster
mtcars |> fgroup_by(cyl) |> fmutate(across(disp:drat, fmean, wt)) |> i()
mtcars |> ftransformv(disp:drat, fmean, cyl, wt, 1, apply = FALSE) |> i()
# ftransform()/fcompute() support list input and ignore attached groupings
mtcars %>% fgroup_by(cyl) %>% ftransform(fselect(., hp:qsec) %>%
           fmedian(TRA = 1) %>% fungroup() %>% fsum(TRA = "/")) |> i()
# Again a faster equivalent: note the use of 'set' to avoid a deep copy
mtcars %>% ftransform(fselect(., hp:qsec) %>% fmedian(cyl, TRA = 1) %>%
                      fsum(TRA = "/", set = TRUE)) %>% i()
# Aggregation: weighted standard deviations
mtcars |> fgroup_by(vs) |> fsummarise(across(disp:drat, fsd, w = wt))
# Grouped linear models (one way of doing it)
qTBL(mtcars) |> fgroup_by(vs) |> fsummarise(reg = list(lm(disp ~ drat)))
# Adding some columns. Use ftransform<- to also replace existing ones
add_vars(iris) <- num_vars(iris) |> fsum(TRA = '%') |> add_stub("perc_")
@
<<echo=FALSE, include=FALSE>>=
iris <- iris2
@
\vspace{-2mm}

@@

% \end{adjustbox}
%}
% \hrrule
\vspace{-5mm}
\textcolor{lightgray}{\hrulefill}\\
{\scriptsize \vspace{-0.5mm}
 Page 1 of 2 \hfill \href{https://creativecommons.org/licenses/by-sa/4.0/}{CC-BY-SA}\ Sebastian Krantz\ \textbullet\ Learn more at \href{https://sebkrantz.github.io/collapse/}{sebkrantz.github.io/collapse}\ \textbullet\ Source code at \href{https://github.com/SebKrantz/collapse}{github.com/SebKrantz/collapse}\ \textbullet\ Updates announced at \href{https://twitter.com/collapse\_R}{twitter.com/collapse\_R} - \#rcollapse\ \textbullet\ Cheatsheet created for \emph{collapse} version 2.0.0\ \textbullet\ Updated: 2023-10
}

@@

\code{any|all[v|NA]}, \code{which[v|NA]}, \code{\%[=|!]=\%}, \code{copyv}, \code{setv}, \code{alloc} \code{missing\_cases}, \code{na\_[insert|rm|omit]}, \code{vlengths}, \code{vtypes}, \code{vgcd}, \code{frange}, \code{fnlevels}, \code{fn[row|col]}, \code{fdim}, \code{seq\_[row|col]}\\
<<eval = FALSE>>=
fsubset(wlddev, year %==% 2010) # 2x faster fsubset(wlddev, year == 2010)
attach(mtcars) # Efficient sub-assignment by reference, various options...
setv(am, 0, vs); setv(am, 1:10, vs); setv(am, 1:10, vs[10:20])
@
% x <- na_insert(rnorm(1e7))
% microbenchmark(na_rm(x), x[!is.na(x)])
% microbenchmark(na.omit(wlddev), na_omit(wlddev, na.attr = TRUE))
% # Efficient subassignment by reference
% with(wlddev, setv(POP, year %==% 2010, fmax(POP)))
% microbenchmark(A = fsubset(wlddev, year == 2010), B = fsubset(wlddev, year %==% 2010))
% x <- rnorm(1e7); xNA <- na_insert(x)
% microbenchmark(setv(xNA, NA, x), )
% with(wlddev, setv(POP, year %==% 2010, fmax(POP)))
% \vspace{-1mm}

\hrrule
\section{Small (Helper) Functions}
\itxt{Functions for (meta-)programming and attributes}\\ [0.5em]
\code{.c}, \code{massign}, \code{\%=\%}, \code{vlabels[<-]}, \code{setLabels}, \code{vclasses}, \code{namlab}, \code{[add|rm]\_stub}, \code{\%!in\%}, \code{ckmatch}, \code{all\_identical}, \code{all\_obj\_equal}, \code{all\_funs}, \code{set[Dim|Row|Col]names}, \code{unattrib}, \code{setAttrib}, \code{copyAttrib}, \code{copyMostAttrib} %, \code{is\_categorical}, \code{is\_date}

@@

\hrrule
\section{API Extensions}
\itxt{Shorthands for frequently used functions}\\ [0.5em]
\code{fselect -> slt, fsubset -> sbt, fmutate -> mtt, [f/set]transform[v] -> [set]tfm[v], fsummarise -> smr,
across -> acr, fgroup\_by -> gby, finteraction -> itn, findex\_by -> iby, findex -> ix, frename -> rnm, get\_vars -> gv, num\_vars -> nv,
add\_vars -> av} \newline

\itxt{Namespace masking}\\ [0.5em]
Can set \code{option(collpse\_mask = c(...))} with a vector of functions starting with f-, to export versions without f-, masking base R or \emph{dplyr}. A few keywords exist to mask multiple functions, see \code{help("collapse-options")}. This allows clean \& fast code, but poses additional namespace challenges:

<<eval = FALSE>>=
# Masking all f- functions and specials n = GRPN and table = qtab
options(collapse_mask = "all")
library(collapse)
# The folowing is 100% collapse code, apart from the base pipe

wlddev |>
  subset(year >= 1990) |>
  group_by(year) |>
  summarise(n = n(), across(PCGDP:GINI, mean, w = POP))

@@

@
The best way to set this option is inside an \code{.Rprofile} file placed in the user or project directory. Use it carefully.

@@

\vspace{-5.5mm}
\textcolor{lightgray}{\hrulefill}\\
{\scriptsize \vspace{-0.5mm}
 Page 2 of 2 \hfill \href{https://creativecommons.org/licenses/by-sa/4.0/}{CC-BY-SA}\ Sebastian Krantz\ \textbullet\ Learn more at \href{https://sebkrantz.github.io/collapse/}{sebkrantz.github.io/collapse}\ \textbullet\ Source code at \href{https://github.com/SebKrantz/collapse}{github.com/SebKrantz/collapse}\ \textbullet\ Updates announced at \href{https://twitter.com/collapse\_R}{twitter.com/collapse\_R} - \#rcollapse\ \textbullet\ Cheatsheet created for \emph{collapse} version 1.8.8\ \textbullet\ Updated: 2022-08
}

>>>>

Calling \code{help("collapse-documentation")} brings up a detailed documentation, which is also available \href{https://fastverse.github.io/collapse/reference/index.html}{online}. See also the \href{https://fastverse.github.io/fastverse/}{\emph{fastverse}} package/project for a recommended set of complimentary packages and easy package management.
%}

@@

\textbf{Examples}
<<>>=
mtcars |> fsubset(mpg > fnth(mpg, 0.95), disp:wt, cylinders = cyl)
mtcars |> colorder(cyl, vs, am, pos = 'after') |> head(2)
i <- base::invisible # These are equivalent, the second option is faster:
mtcars |> fgroup_by(cyl, vs, am) |> fmutate(sum_mpg = fsum(mpg)) |> i()
mtcars |> fmutate(sum_mpg = fsum(mpg, list(cyl, vs, am), TRA = 1)) |> i()
# These are also equivalent (weighted means), again the second is faster
mtcars |> fgroup_by(cyl) |> fmutate(across(disp:drat, fmean, wt)) |> i()
mtcars |> ftransformv(disp:drat, fmean, cyl, wt, 1, apply = FALSE) |> i()
# ftransform()/fcompute() support list input and ignore attached groupings
mtcars %>% fgroup_by(cyl) %>% ftransform(fselect(., hp:qsec) %>%
           fmedian(TRA = 1) %>% fungroup() %>% fsum(TRA = "/")) |> i()
# Again a faster equivalent: note the use of 'set' to avoid a deep copy
mtcars %>% ftransform(fselect(., hp:qsec) %>% fmedian(cyl, TRA = 1) %>%
                      fsum(TRA = "/", set = TRUE)) %>% i()
# Aggregation: weighted standard deviations
mtcars |> fgroup_by(vs) |> fsummarise(across(disp:drat, fsd, w = wt))
# Grouped linear models (one way of doing it)
qTBL(mtcars) |> fgroup_by(vs) |> fsummarise(reg = list(lm(mpg ~ carb)))
# Adding some columns. Use ftransform<- to also replace existing ones
add_vars(iris) <- num_vars(iris) |> fsum(TRA = '%') |> add_stub("perc_")
@
<<echo=FALSE, include=FALSE>>=
iris <- iris2
@
\vspace{-2mm}

@@

% \end{adjustbox}
%}
% \hrrule
\vspace{-5mm}
\textcolor{lightgray}{\hrulefill}\\
{\scriptsize \vspace{-0.5mm}
 Page 1 of 2 \hfill \href{https://creativecommons.org/licenses/by-sa/4.0/}{CC-BY-SA}\ Sebastian Krantz\ \textbullet\ Learn more at \href{https://fastverse.github.io/collapse/}{fastverse.github.io/collapse}\ \textbullet\ Source code at \href{https://github.com/SebKrantz/collapse}{github.com/SebKrantz/collapse}\ \textbullet\ Updates announced at \href{https://twitter.com/collapse\_R}{twitter.com/collapse\_R} - \#rcollapse\ \textbullet\ Cheatsheet created for \emph{collapse} version 2.0.3\ \textbullet\ Updated: 2023-10
}

@@

\code{any|all[v|NA]}, \code{which[v|NA]}, \code{\%[=|!]=\%}, \code{copyv}, \code{setv}, \code{alloc} \code{missing\_cases}, \code{na\_[insert|rm|omit]}, \code{vlengths}, \code{vtypes}, \code{vgcd}, \code{fnlevels}, \code{fn[row|col]}, \code{fdim}, \code{seq\_[row|col]}, \code{vec}\\
<<eval = FALSE>>=
fsubset(wlddev, year %==% 2010) # 2x faster fsubset(wlddev, year == 2010)
attach(mtcars) # Efficient sub-assignment by reference, various options...
setv(am, 0, vs); setv(am, 1:10, vs); setv(am, 1:10, vs[10:20])
@
% x <- na_insert(rnorm(1e7))
% microbenchmark(na_rm(x), x[!is.na(x)])
% microbenchmark(na.omit(wlddev), na_omit(wlddev, na.attr = TRUE))
% # Efficient subassignment by reference
% with(wlddev, setv(POP, year %==% 2010, fmax(POP)))
% microbenchmark(A = fsubset(wlddev, year == 2010), B = fsubset(wlddev, year %==% 2010))
% x <- rnorm(1e7); xNA <- na_insert(x)
% microbenchmark(setv(xNA, NA, x), )
% with(wlddev, setv(POP, year %==% 2010, fmax(POP)))
% \vspace{-1mm}

\hrrule
\section{Small (Helper) Functions}
\itxt{Functions for (meta-)programming and attributes}\\ [0.5em]
\code{.c}, \code{massign}, \code{\%=\%}, \code{vlabels[<-]}, \code{setLabels}, \code{vclasses}, \code{namlab}, \code{[add|rm]\_stub}, \code{all\_identical}, \code{all\_obj\_equal}, \code{all\_funs}, \code{set[Dim|Row|Col]names}, \code{unattrib}, \code{setAttrib}, \code{copyAttrib}, \code{copyMostAttrib}, \code{is\_categorical}, \code{is\_date}

@@

\hrrule
\section{API Extensions and Global Options}
\itxt{Shorthands for frequently used functions}\\ [0.5em]
\code{fselect -> slt, fsubset -> sbt, fmutate -> mtt, [f/set]transform[v] -> [set]tfm[v], fsummarise -> smr,
across -> acr, fgroup\_by -> gby, finteraction -> itn, findex\_by -> iby, findex -> ix, frename -> rnm, get\_vars -> gv, num\_vars -> nv,
add\_vars -> av} \newline

\itxt{Namespace masking and other global options}\\ [0.5em]
Use \code{set\_collpse(mask = c(...))} with a vector of functions starting with f-, to export versions without f-, masking base R and/or \emph{dplyr}. A few keywords exist to mask multiple functions, see \code{help("collapse-options")}. There are also many other global defaults and optimizations that can be controlled with \code{set\_collapse(...)}. Retrieve options using \code{get\_collapse()}.

<<eval = FALSE>>=
# Masking all (f-)functions and changing some defaults (=optimizing)
library(collapse)
set_collapse(mask = "all", na.rm = FALSE, sort = FALSE, nthreads = 4)
# The following is now 100% collapse code and executed without regard for
# missing values, using unsorted grouping and 4 threads (where applicable)
wlddev |>
  subset(year >= 1990 & is.finite(GINI)) |>
  group_by(year) |>
  summarise(n = n(), across(PCGDP:GINI, mean, w = POP))

@@

@

@@

\vspace{-5.5mm}
\textcolor{lightgray}{\hrulefill}\\
{\scriptsize \vspace{-0.5mm}
 Page 2 of 2 \hfill \href{https://creativecommons.org/licenses/by-sa/4.0/}{CC-BY-SA}\ Sebastian Krantz\ \textbullet\ Learn more at \href{https://fastverse.github.io/collapse/}{fastverse.github.io/collapse}\ \textbullet\ Source code at \href{https://github.com/SebKrantz/collapse}{github.com/SebKrantz/collapse}\ \textbullet\ Updates announced at \href{https://twitter.com/collapse\_R}{twitter.com/collapse\_R} - \#rcollapse\ \textbullet\ Cheatsheet created for \emph{collapse} version 2.0.3\ \textbullet\ Updated: 2023-10
}
